Micro Text Mining
========================================================

This is a set of R programs following the tutorials in Mathew Jocker's 'Text Analysis
with R' book.

In this file I go through some of the 'micro' analysis examples. Looking at detailed analyses on one data set.

Let's bring in the whole text for Moby Dick.

```{r ScanMobyDick}
text.v<-scan("http://www.gutenberg.org/cache/epub/2701/pg2701.txt",
what="character", sep="\n")
```
The scan function brings in a text file with one observation per line (/n).

This subsets the data to only include the book text (taking out the editorial stuff etc.).

```{r ExcludeEditorialInfo}
start.v<-which(text.v == "CHAPTER 1. Loomings.")
end.v<-which(text.v == "orphan.")
novel.lines.v<- text.v[start.v:end.v]
````

Collapse data into one string
```{r Collapse}
novel.v<-paste(novel.lines.v, collapse=" ")
````

Turn text to lower case letters and split into words.
```{r Turn2Words}
moby.words.l<-strsplit(tolower(novel.v),"\\W")
#splits the text at every non word character.
moby.words.v<-unlist(moby.words.l)
```

Delete the blanks from the file.
```{r DeleteBlanks}
moby.words.v<-moby.words.v[which(moby.words.v!="")]
````

Calculate the number of times the word 'whale' appears per 100 words of text.
```{r WhaleCount}
# Put a count of the occurecnes of whale into whale.hits.v
whale.hits.v<-length(moby.words.v[which(moby.words.v=="whale")])
# Put a count of total words into total.words.v
total.words.v<-length(moby.words.v)
# now divide
whale.hits.v/total.words.v*100
```

Calc. Number of unique words
```{r Unique}
length(unique(moby.words.v))
#and table by frequency
moby.freqs.t<-table(moby.words.v)
 #Frequency of he/she
moby.freqs.t["he"]/moby.freqs.t["she"]
#and plot top ten words
sorted.moby.freqs.t<-sort(moby.freqs.t , decreasing=T)
sorted.moby.freqs.t[1:10]
plot(sorted.moby.freqs.t[1:10], type="b",
     xlab="Top Ten Words", ylab="Count", xaxt ="n")
axis(1,1:10, labels=names(sorted.moby.freqs.t [1:10]))
```

Word Distribution
------------------
  
Let's look at where certain words appear in the text and create
dispersion plots.

```{r DispersionPlot}
plot(moby.words.v == "ahab", main="Dispersion Plot of `ahab' in Moby Dick",
xlab="Novel Time", ylab="ahab", type="h", ylim=c(0,1), yaxt='n')